---
title: "Data Discovery"
subtitle: "Pluralsight Data Scientist Take-Home Exercise"
author: "Trenton Pulsipher"
date: "`r lubridate::today()`"
output: html_document
---


```{r settings, echo = F, warning = F, message = F, error = F}
knitr::opts_chunk$set(
  echo = F,
  message = F,
  warning = F,
  error = F,
  fig.height = 3,
  fig.width = 9.5,
  cache = F
)

# Libraries 
library(DBI)
library(RSQLite)
library(lubridate)
library(stopwords)
library(tidyverse)
library(trelliscopejs)
library(wordcloud)
library(HSPSUtils) # install_github("HSPS-DataScience/HSPSUtils")
                   # devtools::update_packages("HSPSUtils")
library(rbokeh)
library(ggpubr)
```


```{r dataIn}
# Read in Data

# connect the database created by running .sql file in TablePlus
con <- DBI::dbConnect(RSQLite::SQLite(), "~/Documents/Development/R/pluralsightTest/pluralsightTestDB.sqlite3")

# read in data from the four database tables
question_details <- DBI::dbReadTable(con, "question_details") %>% 
  as_tibble()

question_interactions <- DBI::dbReadTable(con, "question_interactions") %>% 
  as_tibble() %>%
  mutate(date_created = ymd_hms(date_created))

user_assessment_sessions <- DBI::dbReadTable(con, "user_assessment_sessions") %>% 
  as_tibble() %>%
  mutate(
    date_created = ymd_hms(date_created),
    date_modified = ymd_hms(date_modified)
  )

user_interactions <- DBI::dbReadTable(con, "user_interactions") %>% 
  as_tibble() %>%
  mutate(date_created = ymd_hms(date_created))
```


## user_assessment_sessions

Below are 100 randomly selected rows from the dataset. 

```{r dataTable1}
create_datetable <- function(data) {
  ## data table
  DT::datatable(
    data %>%
      sample_n(100),
      options = list(
        pageLength = 5,
        searching = T,
        scrollX = T
      )
    )
}
create_datetable(user_assessment_sessions)
```


### Data Summary

The table below shows several metrics calculated against the various columns/variables. These metrics include: the number of unique values, number of NAs, the maximum value, the minimum value, and the mean/average.
```{r dataSummary1}
create_data_summary <- function(data) {
  # convert column names for use here
  names(data) <- str_replace_all(names(data), "_", ".")

  # Data Summary
  data_summary <- data %>%
    summarise_all(funs(
      numUnique = length(unique(.)),
      nas = sum(is.na(.)),
      max = max(., na.rm = T),
      min = min(., na.rm = T),
      mean = mean(., na.rm = T))) %>%
    gather() %>%
    separate(key, c("key", "stat"), sep = "_") %>%
    spread(key, value)
  # generate datatable from summary
  DT::datatable(
    data_summary,
    options = list(
      pageLength = 5,
      searching = T,
      scrollX = T
    )
  )
}
create_data_summary(user_assessment_sessions)
```


### Categorical Variables in Bar Charts

A quick look at counts of the various levels of the categorical variables found in the data is shown in the following set of bar charts. [Trelliscopejs](https://hafen.github.io/trelliscopejs/) is an R package that enables one to quickly sort and/or filter through various slices (or features) of their previously generated visualizations/plots of their data. We can then sort and filter through the feature set to find anomalies or interesting behaviors that might influence modeling or help us understand basic relationships between variables in the data. 

For this dataset the number of categorical variables is not overwhelming so the power of trelliscope would appear limited. However, when one needs to quickly move through 100s, 1000s or even millions of slices of the data, trelliscope provides a seemless interface to manage so many subsets of data and better discover interesting features.

```{r catBarCharts, fig.height = 4}
create_bar_chart_categorical_trelli <- function(data) {
    data %>%
    # selection of categorical variables isn't automated yet
      select_if(is.character) %>%
      gather_group_by_count() %>%
      ungroup() %>%
    ggplot(aes(x = value, y = Count)) +
      geom_bar(stat = "identity", alpha = 0.5) +
      geom_text(aes(label = scales::comma(Count))) +
      theme_bw() +
      coord_flip() +
      labs(x = "", y = "") +
      facet_wrap(~key, scales = "free")
      # facet_trelliscope(~ key,
      #                   scales = "free",
      #                   self_contained = T,
      #                   width = 600,
      #                   name = "categoricalVariables",
      #                   group = "vars",
      #                   desc = "All Variables of Type Character or Factor")
}
create_bar_chart_categorical_trelli(user_assessment_sessions)
```


### Numeric Variables in Histograms

```{r numHistograms1, fig.height = 7}
create_hist_numeric <- function(data) {
  ## histogram of all numeric variables
  data %>%
    select(-ends_with("id")) %>%
    select_if(is.numeric) %>%
    gather() %>%
    filter(value >= 0) %>%
  ggplot(aes(x = value)) +
    geom_histogram(bins = 30) +
    scale_y_continuous(labels = scales::comma) +
    # scale_x_log10(breaks = c(0.1, 1, 10, 100, 1000, 10000),
    #               labels = c(0.1, 1, 10, 100, 1000, 10000)) +
    facet_wrap(~ key, scales = "free", ncol = 3) +
    theme_bw() +
    labs(x = "", y = "Count")
}
create_hist_numeric(user_assessment_sessions)
```


### Time Series Variables

As noted in the section below, outliers and NAs make it difficult to initially see the trends over time of these two time series related variables.
```{r ts1, fig.height = 4, fig.width = 8}
create_time_series <- function(data, title) {
  # time series
  data %>%
    mutate(ymd = ymd(as.Date(date))) %>%
    group_by(ymd) %>%
    summarise(Count = n()) %>%
    ungroup() %>%
    figure(xlab = "", title = title) %>%
      ly_points(x = ymd, y = Count,
               hover = list(
                 Date = ymd, Count = Count
               ))
}
create_time_series(user_assessment_sessions %>% select(date_created) %>% rename(date = date_created), 
                   title = "user_assessment_sessions date_created")
```

```{r ts2, fig.height = 4, fig.width = 8}
create_time_series(user_assessment_sessions %>% select(date_modified) %>% rename(date = date_modified), 
                   title = "user_interactions date_modified")
```

